### 2016-02-09 Manuscript/BiologyLetters/medianSummary.tex
Media summary

\documentclass{article} % Setting the class of the document

%For word count, you can install the super cool (and easy to use) WordCount package through Package Control

\begin{document} % starting the document

%What's the story
To understand changes of biodiversity, it is crucial to combine both living and fossil species in the same analysis.
We can do so by using the DNA from living species and enough anatomical data in common between living and fossil species.
%What we found
In this study we found that only a small amount of living mammals have available anatomical data and that this could be a problem for combining them with fossils.
%How's that gonna solve the world
We suggest that an increase in sampling effort in natural history museum collections, where this data lies, is necessary to efficiently combine living and fossil species in the same analysis.

\end{document} % Ending the document (don't forget this one!)