### 2016-02-12 Manuscript/BiologyLetters/medianSummary.tex
My attempt
Feel free to revert, I tried to make it more media friendly

\documentclass{article} % Setting the class of the document

%For word count, you can install the super cool (and easy to use) WordCount package through Package Control

\begin{document} % starting the document

Scientists use family trees to understand changes in biodiversity through time. But the trees often only use living species, ignoring extinct groups like mammoths. Therefore we must place living and 
extinct species in the same trees. We can do so using the DNA from living species and anatomical data from living and extinct species. However, we found that few living mammals have enough available anatomical data to determine how they are related to extinct species. We suggest that we must collect more anatomical data from living species in museum collections to efficiently create family trees of living and extinct species.

\end{document} % Ending the document (don't forget this one!)